---
title: "Lab 01 Notebook"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This is the replication notebook for Lab-01 from the course SOC 4650/5650: Introduction to GISc.

## Project Set Up

```{r setup}
knitr::opts_knit$set(root.dir = here::here())
```

## Load Dependencies
The following code loads the package dependencies for our analysis:

```{r package-load}
library(ggplot2) # creating plots
```

## Load Data
The following code loads the data package and assigns our data to a data frame in our global environment:

```{r data-load}
library(stlData) # data source
medianInc <- stlIncome
```

## Part 1
### Question 10
The first question in this assignment asks for a histogram of the `pctElevated` variable.

```{r income-plot-01}
ggplot() +
  geom_histogram(medianInc, mapping = aes(mi15))
```

### Question 11
Next, we want to save our plot using the `ggsave()` function:

```{r}
ggsave("results/incomeHistogram.png")
```
